---
title: "BIMM 143 Class 17"
author: "Anika Bhattacharjya (A15459876)"
date: "11/23/2021"
output: 
  pdf_document: 
    latex_engine: xelatex
---

# Getting Started

```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
#head(vax)
```
> Q1. What column details the total number of people fully vaccinated?

The 9th column that says "persons_fully_vaccinated."

> Q2. What column details the Zip code tabulation area? 

The 2nd column labeled "zip_code_tabulation_data."

```{r}
library(lubridate)
```
```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

> Q3. What is the earliest date in this dataset? 

```{r}
vax$as_of_date[1]
```

> Q4. What is the latest date in this dataset? 

```{r}
vax$as_of_date[nrow(vax)]
```

```{r}
library(devtools)
```
```{r}
library(skimr)
```
```{r}
skimr::skim(vax)
```
> Q5. How many numeric columns are in this dataset?

9

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

```{r}
sum( is.na(vax$persons_fully_vaccinated) )
```

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

```{r}
library(plyr)
```
```{r}
signif(sum( is.na(vax$persons_fully_vaccinated) )/81144*100,2)
```

> Q8. [Optional]: Why might this data be missing?

It may be missing because of human record keeping error or no one reporting vaccinations.

# Working with dates

```{r}
library(lubridate)
```
```{r}
today()
```

Look at the as_of_date column
```{r}
# Speciffy that we are using the Year-mont-day format
vax$as_of_date <- ymd(vax$as_of_date)
```
```{r}
today() - vax$as_of_date[1]
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

> Q9. How many days have passed since the last update of the dataset?

```{r}
today() - vax$as_of_date[nrow(vax)]
```

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(unique(vax$as_of_date))
```

# Working with Zip Codes

"zipcodeR" wouldn't work for me so Professor Grant said to skip this section.

# Focus on San Diego Area
```{r}
# Subset to San Diego county only areas
sd <- vax[ vax$county == "San Diego" , ]
```
```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```
```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

> Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
length(unique(sd$zip_code_tabulation_area))
```

> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
which.max(sd$age12_plus_population)
```
```{r}
sd$zip_code_tabulation_area[23]
```

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?

```{r}
sd.vax <- filter(vax, county == "San Diego" &
              as_of_date == "2021-11-09")
```

```{r}
mean(sd.vax$percent_of_population_fully_vaccinated, na.rm=TRUE)
```


> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

```{r}
library(ggplot2)
```
```{r}
?ggplot
```

```{r}
ggplot(sd.vax) + geom_histogram(aes(x=percent_of_population_fully_vaccinated)) + labs(x = "Percent Fully Vaccinated as of 2021-11-09", y = "Frequency")
  
```
# Focus on UCSD/La Jolla
```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```
> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area

```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
```
# Comparing 92037 to other similar sized areas

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

```

> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) + geom_hline(yintercept = 0.66, col = "red") +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
quantile(vax.36$percent_of_population_fully_vaccinated)
```
```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```

> Q18. Using ggplot generate a histogram of this data

```{r}
ggplot(vax.36) + geom_histogram(aes(x=percent_of_population_fully_vaccinated)) + labs(x = "Percent Fully Vaccinated", y = "Count")
```

> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```
```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
92040 is below average (0.6629812) and 92109 is above.

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(0,1) +
  labs(x= "Date", y= "Percent Vaccinated",
       title= "Vaccination rate across California",
       subtitle="Only areas with a population above 36k are shown") +
  geom_hline(yintercept = 0.66, linetype= "dashed")
```

> Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?

I think it would be better to do it virtually since people won't have time to get properly tested by Tuesday if they come back on Sunday.