---
title: "Class_08"
author: "Anika Bhattacharjya (A15459876)"
date: "10/21/2021"
output: github_document
---

First up is clustering methods

#Kmeans clustering

The function in base R to do Kmeans clustering is called 'kmeans()'

First make up some data where we know what the answer should be:

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
x
plot(x)
```

> Q. Can we use kmeans to cluster this data setting k 2 and nstart to 20

```{r}
km <- kmeans(x, centers=2, nstart=20)
km
```

> Q. How many points are in each cluster?

```{r}
km$size
```

> Q. What 'component' of your result object details cluster assignment/membership?

```{r}
km$cluster
```

> What 'component' of your result object details cluster centers?

```{r}
km$centers
```


> Q. Plot x colored by the kmeans cluster assignment and add cluser centers as blue points

```{r}
plot(x, col= km$cluster)
points(km$centers, col="blue", pch=15, cex=2)
```

Analyze same data with hclust()

```{r}
hc <- hclust (dist(x))
hc
```

There is a plot method for hclust result object
```{r}
plot(hc)
```

To get our cluster membership vector we need to do a little more work. We have to "cut" the tree where we think it makes sense. For this we use `cutree()` function.

```{r}
cutree(hc, h = 6)
```
 You can also call `cutree()` setting k = numner of grps/clusters you want.
 
```{r}
cutree(hc, k=2)
```
 
 Make our results plot
 
```{r}
grps <- cutree(hc,k=2)
plot(x, col=grps)
```
 
#PCA of UK food data

> import data

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
```

> Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?

```{r}
dim(x)
```

> Checking your data

Preview first 6 rows

```{r}
x
```

Row-names are incorrectly set as the first column of our x data fram. Use minus indexing as one option.
```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```
 There is a problem here because if we keep inputing the function it takes a column away each time.

Another method:
```{r}
x <- read.csv(url, row.names=1)
head (x)
```
```{r}
dim(x)
```

Now we have correct dimensions
> Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

The second option is the better option because it doesn't keep taking away columns.

Plot the data 
```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

> Q3: Changing what optional argument in the above barplot() function results in the following plot?

Set beside to false
```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

> Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?

```{r}
pairs(x, col=rainbow(10), pch=16)
```

> Q6. What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set?

Can't really tell the difference at this point the way the data is presented.

Use PCA Funtion
```{r}
# Use the prcomp() PCA function
pca <- prcomp(t(x))
summary(pca)
```

> Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points.

```{r}
#Plot PC1 vs PC2
colors <- c("red", "Orange", "Green", "Blue")
plot(pca$x[,1], pca$x[,2])
text(pca$x[,1], pca$x[,2], colnames(x), col= colors)
```

Variable Loadings!

```{r}
# Focus on PC1 as it accounts for > 90% of variance
par(mar=c(10, 3, 0.35, 0))
barplot(pca$rotation[,1], las=2)
```

